### 2020-06-23 appendix.tex

@@

>>>>

Any $\bmd$-majorization problem can be rephrased as a majorization problem in a higher dimensional space due to the embedding map.
\begin{definition}
    The embedding map $\Gamma_{\bmd}:\mathbb{R}^n \mapsto \mathbb{R}^D, D = \sum\limits_{i=1}^n d_i$ is the function
    \begin{equation}
        \Gamma_{\bmd}(\bmz) = \bigoplus_{i=1}^n z_i\  \frac{1}{d_i}\bm{1},
    \end{equation}
    where $\bm{1}/d_i$ is the $d_i$-dimensional uniform distribution.
    The left inverse $\Gamma_{\bmd}^{-1}: \mathbb{R}^D \mapsto \mathbb{R}^n$ is defined to sum up the elements in each block of $\Gamma_{\bmd}(\bmz)$, so that
    \begin{equation}
        \Gamma_{\bmd}^{-1}(\oplus_{i=1}^n z_i \bm{1}/d_i) = \bmz.
    \end{equation}
    This is not a right inverse, because $\Gamma_{\bmd}$ is not surjective.
\end{definition}
The direct sum simply amounts to listing the uniform distributions one after the other.
The embedding map maps the Gibbs distribution to the uniform distribution, $\Gamma_{\bmd}(\bmd) = \bm{1}/D$.
Then, a non-increasing ordering $\Gamma_{\bmd}(\bmz)^\downarrow$ in the new space, corresponds to the so-called ``$\beta$-ordering'' of the original vector denoted by the permutation $\pi$ mapping $(z_i/d_i) \mapsto (z_i/d_i)^\downarrow$ for all $i=1,\dots,n$.
Intuitively, $\beta$-ordering tends to simply non-increasing ordering as $\beta$ tends to 0.
This ordering naturally leads to a generalised notion of the Lorenz curve.

@@

The Lorenz curve can be explicitly expressed as
    \begin{equation}
		\lc{\rho}{\sigma}(x) = (\bmu^\downarrow)_i x + \lc{\rho}{\sigma}(i) - (\bmu^\downarrow)_i x_i,\ x \in (x_{i-1}, x_i],
	\end{equation}
	for all $i=1,\dots,d$ with $x_0 \coloneqq 0$.

### 2020-06-23 magic_paper.tex

\ddd{[Things we must emphasize:
\begin{enumerate}
\item We can tackle more `physicsy' questions like: how much magic can be distilled via free operations with some given fixed-point structure? E.g. 
\item This allows a diagnostic on the kind of operations needed to do good distillation. I.e. what fixed point structure should they have?
\item We go beyond magic monotones and replace a monotone with a $\sigma$-fragment.
\item We can get both upper and lower bounds on magic distillation.
\end{enumerate}•
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Majorization of quasi-probabilities in the $\sigma$--fragments}\label{sec:major}

@@

\begin{theorem}\label{thm:sigmamajor}
    Let $\R = (\F, \O)$ be a theory of magic. Then $\rho \longrightarrow \tau$ is possible then  $\W{\tau} \prec_\sigma \W{\rho}$ within a $\sigma$--fragment.
\end{theorem}
\begin{proof}Suppose we can convert $\rho$ into $\tau$ in the magic theory. Thus there is some $\O_\sigma$ and some $\E \in \O_\sigma$ such that $\E(\rho) = \tau$, and $\E(\sigma)=\sigma$. Therefore, the Wigner distribution of this free operation satisfies $\W{\E} \in \stochw$ and $\W{\E}\W{\rho} = \W{\tau}$. Since $\sigma$ is full-rank and free, its Wigner distribution is strictly positive in all components, so it directly follows from~\cref{def:dmajor} that $\W{\tau} \prec_{\sigma} \W{\rho}$.
\end{proof}
This result can be understood as an extension of the idea of a magic monotone, where we replace $\M(\tau) \le \M(\rho)$ with $W_\tau \prec_\sigma W_\rho$. However the majorization constraints can be used to place upper bounds on magic state distillation in a way that allows one to incorporate the physics of the allowed operations -- this enables us to bound how much magic can be distilled via quantum operations that, for example, preserve the equilibrium state of the system, or via operations that are symmetric about the $Z$-axis of the Bloch sphere.

This approach can also provide \emph{lower bounds} on distillation, however now more structure about the specific free operations must be included. This is more involved and we discuss this later in the paper.

\subsection{Majorization computations in a $\sigma$--fragment}
A numerically efficient reformulation of $\bmd$-majorization is provided by Lorenz curves.
Let the vector $\bmu^\downarrow$ denote the vector $\bmu \in \reals^d$ with its components arranged in non-increasing order.
\begin{definition}[\textbf{Lorenz curve}]\label{def:lc}
    Let $\bmz, \bmd \in \reals^d$, where the components of $\bmd$ are positive with $D = \sum_{i=1}^d d_i$ and denote by $\bmu \coloneqq (z_1/d_1, \dots, z_d/d_d)^T$ the vector of ratios between the corresponding components of $\bmz$ and $\bmd$.

    Finally, denote by $\pi: \zd \mapsto \zd$ the permutation that sorts $\bmu$, $(\bmu^\downarrow)_i = u_{\pi(i)}$ for all $i=1,\dots,d$.

    Consider the piecewise linear curve obtained by joining the points $\{(0,0)\} \cup \{ (x_k, \lc{\bmz}{\bmd}(k)) \}_{k=1,\dots,d}$, where
    \begin{equation}\label{eq:lorenz}
        (x_k, \lc{\bmz}{\bmd}(k)) \coloneqq \left( \frac{1}{D}\sum_{i=1}^k d_{\pi(i)}, \sum_{i=1}^k z_{\pi(i)} \right) \in \mathbb{R}^2.
    \end{equation}
    The \emph{Lorenz curve} of vector $\bmz$ with respect to $\bmd$ is then $\lc{\bmz}{\bmd}(x)$ for all $x \in [0,1]$

    Expressed explicitly,
    \begin{equation}
		\lc{\rho}{\sigma}(x) = (\bmu^\downarrow)_i x + \lc{\rho}{\sigma}(i) - (\bmu^\downarrow)_i x_i,\ x \in (x_{i-1}, x_i],
	\end{equation}
	for all $i=1,\dots,d$ with $x_0 \coloneqq 0$.
\end{definition}
\nick{Move the complicated last equation to appendix?} \ddd{[Yes. Just describe in words, is simpler. the reader with being able to understand how to construct a piecewise linear function.]}
Components $x_k$ are rescaled by $D$ so that comparison of curves with unequal dimensions is possible.
In fact, the Lorenz curves $\lc{\bmz}{\bmd}$ and $\lc{\bmz \otimes \bmd}{\bmd \otimes \bmd}$, where $\otimes$ denotes the Kronecker product, are identical.
Furthermore, a Lorenz curve $\lc{\bmz}{\bmd}(x)$ is always concave in $x$, since it consists of $d$ line segments each with slope $(\bmu^\downarrow)_i$ for $i=1,\dots,d$ which is non-increasing by definition.

@@

We first show that the maximum of $\lc{\rho}{\sigma}$ is independent of the $\sigma$--fragment and directly related to the state sum-negativity $\sn{\rho}$ discussed in~\cref{sec:mono}
\begin{lemma}\label{lem:lcmax}
	Given a quantum state $\rho$, the maximum of its Lorenz curve $\lc{\rho}{\sigma}$ is independent of the $\sigma$--fragment in which it is defined and equal to $1+\sn{\rho}$.
\end{lemma}
\begin{proof}
	Borrowing the notation of~\cref{def:lc}, the derivative of the Lorenz curve $\lc{\rho}{\sigma}$ is
	\begin{equation}
		\frac{\rm{d}}{\rm{d}x} \lc{\rho}{\sigma}(x) = (\bmu^\downarrow)_i,\ x \in (x_{i-1}, x_i],
	\end{equation}
\ddd{[The function is not differentiable at the key points involved. The maximum is obtained when you sum all the non-negative components.]}

	Note that all components of $\W{\sigma}$ are positive, so $(\bmu^\downarrow)_i \geq 0$ if and only if $(\W{\rho})_i \geq 0$ for any $i=1,\dots,d^2$.

@@

\subsection{Distilling out magic in a fragment}
Consider, as an example, the problem of purifying out magic from a noisy state of the form
\begin{equation}
    \rho_{\rm{S}}(\epsilon) = (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \sigma,
\end{equation}
in $\O_\sigma$.
At noise levels $\epsilon \geq \frac{3}{4}$, the state is sufficiently noisy so as the Wigner distribution $\W{\rho_{\rm{S}}(\epsilon)}$ no longer contains negativities. 
At any $\epsilon \leq \frac{3}{4}$, we can investigate the distillation process resulting in a single-copy state with sufficiently low target $\epsilon_{\tau}$,
\begin{equation}\label{eq:sdist}
	\rho_{\rm{S}}(\epsilon)^{\otimes k} \xrightarrow{\E \in \O_\sigma} \rho_{\rm{S}}(\epsilon_{\tau})
\end{equation}

>>>>

\ddd{Things we must emphasize:
\begin{enumerate}
\item We can tackle more `physicsy' questions like: how much magic can be distilled via free operations with some given fixed-point structure? E.g. 
\item This allows a diagnostic on the kind of operations needed to do good distillation. I.e. what fixed point structure should they have?
\item We go beyond magic monotones and replace a monotone with a $\sigma$-fragment.
\item We can get both upper and lower bounds on magic distillation.
\end{enumerate}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Majorization of quasi-probabilities in $\sigma$--fragments}\label{sec:major}

@@

\begin{theorem}\label{thm:sigmamajor}
    Let $\R = (\F, \O)$ be a theory of magic. Then $\rho \longrightarrow \tau$ is possible then  $\W{\tau} \prec_\sigma \W{\rho}$ within a $\sigma$--fragment.
    \nick{I understand that dropping the full-rank assumption on $\sigma$ makes the statement a lot stronger, but also incorrect e.g. in the case of the replacement channel $\E(\rho) = \ketbra{0}$, since $d$-majorization in $\O_{\ketbra{0}}$ is not well-defined.}
\end{theorem}
\begin{proof}
Suppose we can convert $\rho$ into $\tau$ in the magic theory. 
Thus there is some $\O_\sigma$ and some $\E \in \O_\sigma$ such that $\E(\rho) = \tau$, and $\E(\sigma)=\sigma$. 
Therefore, the Wigner distribution of this free operation satisfies $\W{\E} \in \stochw$ and $\W{\E}\W{\rho} = \W{\tau}$. 
Since $\sigma$ is full-rank and free, its Wigner distribution is strictly positive in all components, so it directly follows from~\cref{def:dmajor} that $\W{\tau} \prec_{\sigma} \W{\rho}$.
\end{proof}
This result can be understood as an extension of the idea of a magic monotone, where we replace $\M(\tau) \leq \M(\rho)$ with $\W{\tau} \prec_\sigma \W{\rho}$. 
However the majorization constraints can be used to place upper bounds on magic state distillation in a way that allows one to incorporate the physics of the allowed operations -- this enables us to bound how much magic can be distilled via quantum operations that, for example, preserve the equilibrium state of the system, or via operations that are symmetric about the $Z$-axis of the Bloch sphere.

This approach can also provide \emph{lower bounds} on distillation, however now more structure about the specific free operations must be included. 
\nick{This is more involved and we discuss this later in the paper.
Also mention relative majorization somewhere}

\subsection{Majorization computations in $\sigma$--fragments}
A numerically efficient reformulation of $\bmd$-majorization is provided by Lorenz curves.
Let the vector $\bmu^\downarrow$ denote the vector $\bmu \in \reals^d$ with its components arranged in non-increasing order.
\begin{definition}[\textbf{Lorenz curve}]\label{def:lc}
    Let $\bmz, \bmd \in \reals^d$, where the components of $\bmd$ are positive with $D = \sum_{i=1}^d d_i$ and denote by $\tilde{\bmz} \coloneqq (z_1/d_1, \dots, z_d/d_d)^T$ the vector of ratios between the corresponding components of $\bmz$ and $\bmd$.

    Finally, denote by $\pi: \zd \mapsto \zd$ the permutation that sorts $\tilde{\bmz}$, $(\tilde{\bmz}^\downarrow)_i = z_{\pi(i)}$ for all $i=1,\dots,d$.

    Consider the piecewise linear curve obtained by joining the points $\{(0,0)\} \cup \{ (x_k, \lc{\bmz}{\bmd}(k)) \}_{k=1,\dots,d}$, where
    \begin{equation}\label{eq:lorenz}
        (x_k, \lc{\bmz}{\bmd}(k)) \coloneqq \left( \frac{1}{D}\sum_{i=1}^k d_{\pi(i)}, \sum_{i=1}^k z_{\pi(i)} \right) \in \mathbb{R}^2.
    \end{equation}
    We define the set of points on this curve, $\lc{\bmz}{\bmd}(x),\ x \in [0,1]$, as the \emph{Lorenz curve} of vector $\bmz$ with respect to $\bmd$.
\end{definition}
Components $x_k$ are rescaled by $D$ so that comparison of curves with unequal dimensions is possible.
In fact, the Lorenz curves $\lc{\bmz}{\bmd}$ and $\lc{\bmz \otimes \bmd}{\bmd \otimes \bmd}$, where $\otimes$ denotes the Kronecker product, are identical.
Furthermore, a Lorenz curve $\lc{\bmz}{\bmd}(x)$ is always concave in $x$, since it consists of $d$ line segments each with slope $(\tilde{\bmz}^\downarrow)_i$ for $i=1,\dots,d$ which is non-increasing by definition.

@@

We first show that the maximum of $\lc{\rho}{\sigma}$ is independent of the $\sigma$--fragment and directly related to the state sum-negativity $\sn{\rho}$ discussed in~\cref{sec:mono}
\begin{lemma}\label{lem:lcmax}
	Given a quantum state $\rho$, the maximum of its Lorenz curve $\lc{\rho}{\sigma}$ is independent of the $\sigma$--fragment in which it is defined and equal to $1+\sn{\rho}$.
\end{lemma}
\begin{proof}
	Borrowing notation from~\cref{def:lc}, let $k_\star$ be the index of the smallest non-negative component $\tilde{\bmz}_{\pi(i)}$ of $\tilde{\bmz}^\downarrow$.
	It is clear by the construction of the Lorenz curve described in~\cref{def:lc} that the maximum is achieved at
	\begin{equation}
		x_{k_\star} \coloneqq \sum_{i=1}^{k_\star} (\W{\sigma})_{\pi(i)},
	\end{equation}
	and takes the value
	Note that all components of $\W{\sigma}$ are positive, so $(\bmu^\downarrow)_i \geq 0$ if and only if $(\W{\rho})_i \geq 0$ for any $i=1,\dots,d^2$.

@@

\subsection{Distilling out magic in a fragment}
\nick{why this new subsection?}
Consider, as an example, the problem of purifying out magic from a noisy state of the form
\begin{equation}
    \rho_{\rm{S}}(\epsilon) = (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \sigma,
\end{equation}
in $\O_\sigma$.
At noise levels $\epsilon \geq \frac{3}{4}$, the state is sufficiently noisy so as the Wigner distribution $\W{\rho_{\rm{S}}(\epsilon)}$ no longer contains negativities. 
At any $\epsilon \leq \frac{3}{4}$, we can investigate the distillation process resulting in a single-copy state with sufficiently low target $\epsilon_{\tau}$,
\begin{equation}\label{eq:sdist}
	\rho_{\rm{S}}(\epsilon)^{\otimes k} \xrightarrow{\E \in \O_\sigma} \rho_{\rm{S}}(\epsilon_{\tau})
\end{equation}
